---
title: "comm_figures_2"
author: "Brian Kennedy"
date: "March 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
library(ggrepel)
library(viridis)
library(tidyverse)
library(cowplot)
library(ggthemes)
library(extrafont)
library(treemapify)

options(scipen = 999) 

# directory for r scripts / functions
code.path <- "C:/Users/brian/Dropbox/COCA--diversity/Community data/scripts/"

# directory for datasets
data.path <- "C:/Users/brian/Dropbox/COCA/DATA/GMRI_aggregated data/"

# directory for datasets
image.path <- "C:/Users/brian/Dropbox/COCA--diversity/Community data/figures/"

# directory for community report datasets (need to merge)
commdata.path <-  "C:/Users/brian/Dropbox/COCA--diversity/Community data/data/"

geo.table <- as_tibble(read.csv(paste(data.path,"geo_table.csv", sep = "")))

port.geocode <- as_tibble(read.csv(paste(commdata.path,"data_geocoded_portnames.csv", sep = ""))) 

# reading in raw landings data and joining with the geo dataset / need to change out to actual raw landings

raw.landings <- as_tibble(read.csv(paste(data.path,"byport_thin.csv", sep = ""))) %>%
  left_join(., geo.table, by = "PORT_CODE") %>% drop_na(port_tidy) 

raw.landings.trimmed <- raw.landings %>% 
  dplyr::select(YEAR, SPECIES, value, port_tidy) %>% 
  na.omit()

# reading in function that creates time series w/ top n species 

source(paste(code.path, "topp_spp_function.R", sep = ""))

# reading in sheet with nicely formatted names for species and ports
nice.names <- read.csv(paste(commdata.path,"nice_names.csv", sep=""))

# avg. values for baseline period
portland <- top_spp(raw.landings.trimmed, "PORTLAND_ME", 4, 2011,2015) %>% 
  left_join(., nice.names, by = "spp.top") %>% 
  filter(between(yr.column, 2011,2015)) %>% 
  group_by(spp.nice) %>% 
    summarise(
      avg.value = mean(value)) %>% 
          ungroup() %>% 
        mutate(total.avg.value = sum(avg.value),
               perc.value = avg.value / total.avg.value) %>% 
          coord.barz(.) %>%  #  calling function to make label coordingates
            mutate(y.label = paste(round(y*100, digits = 0),"%")) 

# total annual landings for all species 
portland.total <- top_spp(raw.landings.trimmed, "PORTLAND_ME", 5, 2011,2015) %>% 
  group_by(yr.column) %>% 
  summarise(
    total.value = sum(value))


# MAP DATA the byport landings dataset
map.landings <- raw.landings %>% 
  filter(between(YEAR, 2011,2015)) %>% na.omit(port_tidy) %>% 
  left_join(.,nice.names, by = c("port_tidy" = "port.tidy")) %>% 
  mutate(
    county_pad =  sprintf("%03d", CountyFIPS),  #inserting 0 in COUNTY
    fips_geo = as.numeric(paste0(StateFIPS,county_pad))) %>%  #joinable w/ map
  group_by(fips_geo) %>% 
  summarise(
    value = mean(value))

# reading in county fips names from ggmap
countyfipscodes <- as_tibble(maps::county.fips) %>% 
  separate(polyname, c("state", "county"), sep = ",")

# reading in cnty shapefile from ggmap
cnty <- as_tibble(map_data("county"))

# merging the fips names with the shape file (county names by county names)
map_cnty <- left_join(cnty, countyfipscodes, by = c("region" = "state", "subregion" = "county")) %>% 
  rename( gg_state = region,  gg_cnty = subregion)

# merging the cnty shapefile and fips names WITH our landings dataset,
#excluding counties w/ no landings
cnty_input_data <- left_join(map_cnty, map.landings, by = c("fips" = "fips_geo")) %>% 
  drop_na(value)

# state base layer shapefile, trimmed to relevant states
state_data <- as_tibble(map_data("state")) %>% 
  rename(state_group = group, state_order = order) %>% 
  filter(region %in% c("maine", "new hampshire", "massachusetts", 
                       "vermont", "rhode island", "connecticut", "new york", "new jersey",
                       "pennsylvania", "delaware", "maryland", "district of columbia", "virginia", 
                       "north carolina"))

#create dataset for points 
coca_port_list <- port.geocode %>% 
  left_join(., nice.names, by = c("port_tidy" = "port.tidy")) %>% 
      distinct(lon, lat, port.nice) %>% drop_na(port.nice)


gear_spp_input <- read.csv("C:/Users/brian/Dropbox/COCA--diversity/Community data/data/gear_spp_input.csv")

just.gear <- gear_spp_input %>% group_by(gear_type) %>% 
  summarise(value = sum(value))

```

## Goal: 

Produce figures for page 1 of the final report. The figures should have consistent formatting and allow for the right 'flow' of information as the reader progresses across the page. 

## Formatting Considerations

*  Text: All figures are *roboto condensed* 
*  Palette: I tried to implement the GMRI color scale as per Elijah. At this stage this just means a continuous or discrete palette between gray and blue. Saturation / transparency was hard to dial and varies across the document 
*  Borders: I tried to go with a white borders on major plot elements


## Landings

Simple area chart of annual landings w/ the baseline period treated with a distinct fill. 

**Data**: LANPORT / GAR 82_2015 


```{r landings trend, echo=FALSE, warning=FALSE}
ggplot(data = portland.total, 
                             aes(yr.column, total.value/1000000)) +
  geom_area(fill = gmri.gray, alpha = .6, colour = "white", size = 0.1) +
  geom_area(data = portland.total %>% filter(yr.column > 2010),
            fill = gmri.blue, alpha = .8, colour = "white", size = 1.05) +
  geom_line(colour = "white", size = .8) +
  scale_x_continuous(limits = c(1982, 2020), breaks = seq(min(portland.total$yr.column), 2015, by = 8)) +
  scale_y_continuous(labels = scales::dollar) +
  ylab("Millions of $") + 
 # ggtitle("Annual Landings Value",
  #        subtitle = "PORT OF INTEREST: 1982 - 2015") + 
  theme_tufte() + 
  theme(axis.title.x = element_blank(),
        text = element_text(family = "Roboto Condensed"),
        aspect.ratio = .2)  
```

## Top Species

Bar Plot of the 'Top 4' species + 'Other' 

**Data**: LANPORT / GAR 82_2015, this can be changed to CFDERS easily 


```{r bar plot, echo=FALSE, warning=FALSE}

ggplot() + 
  geom_bar(data = portland, aes(x =1, y = perc.value, fill = 
                                  fct_reorder( spp.nice , avg.value, .desc = TRUE)), stat = "identity", 
           colour = "white", size = 1.05, alpha = .6) +  
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(limits = c(.5,1.5)) + 
  coord_flip() +
  scale_fill_brewer(palette(mypalette)) +
  geom_text(data = portland, aes(x = x, y = y, label = y.label),
             colour = "black", check_overlap = TRUE) +
  guides(fill=guide_legend(title="Top Species", nrow = 2)) + 
  theme(axis.title = element_blank(),
        axis.line = element_blank(), axis.ticks = element_blank(),
        axis.text = element_blank(),legend.position = "bottom",
        text = element_text(size = 8, family = "Roboto Condensed"),
        aspect.ratio = .2)

```


## Treemaps {.tabset}

### Gear and Species

Tree map built from avg. landings value across the baseline period. I have concerns that the species sub-group is TMI and complicates our bar chart.

**Data**: CFDERS 2011 - 2015 


```{r treemap gear spp, echo=FALSE, warning=FALSE}
ggplot(gear_spp_input, aes(area = value / 1000, label = spp_top, fill =value,
                           subgroup = gear_type)) +
    geom_treemap(alpha = .8, colour = "white") +
  geom_treemap_text(colour = "white", place = "center", reflow = T, alpha = .6) +
 geom_treemap_subgroup_border(colour = "white", alpha = .8) + 
 geom_treemap_subgroup_text(place = "topleft", alpha = 0.8, colour =
                               "black") +
  scale_fill_gradient(low = gmri.blue, high = gmri.gray, labels = scales::dollar) +
  labs(fill = "Landings Value (Thousands of $)") +
#  ggtitle("Avg. Landings Value by Gear Type and Species",
 #         subtitle = "Computed using 2011 - 2015 Values / Area Proportional to Value") +
  theme_tufte() + 
  theme(legend.position = "bottom",
        text = element_text(family = "Roboto Condensed"))
```

### Just Gear 

Tree map built from avg. landings value across the baseline period. Species subgroup is excluded here. 

**Data**: CFDERS 2011 - 2015 

```{r treemap just gear, echo=FALSE, warning=FALSE}

ggplot(gear_spp_input %>% ungroup() %>% group_by(gear_type) %>% 
         summarise(value = sum(value)/10), aes(area = value, label = gear_type, fill =value )) + 
  geom_treemap(alpha = .8, colour = "white") +
  geom_treemap_text(colour = "white") + 
  scale_fill_gradient(low = gmri.blue, high = gmri.gray, labels = scales::dollar) +
  guides(fill=guide_legend(title="Share of Landings")) +
  theme_tufte() + 
  theme(legend.position = "bottom",
        text = element_text(family = "Roboto Condensed"),
        aspect.ratio = .8)

```

## Scatterplot 

Scatter plot with catch diversity and log transformed Avg. Value. I have excluded axis values here. 

```{r scatterplot, echo=FALSE, warning=FALSE}

value.div <-  port.value.div(raw.landings) %>% filter(avg.total.value > 1000000)

data.filt <- value.div %>% filter(port_tidy == "PORTLAND_ME") %>% 
    left_join(., nice.names, by = c("port_tidy" = "port.tidy"))

ggplot() + 
  geom_point(data = value.div, aes(x = log(avg.total.value), y = avg.index),
             color = gmri.gray, alpha = .6, size = 1) +
  geom_point(data = data.filt, aes(x = log(avg.total.value), y = avg.index),
             color = gmri.blue, size = 2.5) +
  geom_text(data = data.filt, aes(x = log(avg.total.value), y = avg.index, label = port.nice),
            hjust = 0, nudge_x = .13) + 
  ylab("Catch Diversity") +
  xlab("Avg. Total Value ") +
    theme(aspect.ratio = .26,
          text = element_text(size = 10, family = "Roboto Condensed"),
          axis.ticks = element_blank(), axis.text = element_blank())


```



## Map

Basic baseman. Palette has been reversed. 

**Data**:  LANPORT / GAR 82_2015 + default GGMAP categories 

```{r map, echo=FALSE, warning=FALSE}
ggplot() +
  geom_polygon(data = state_data, aes(x=long, y=lat, group=state_group),
               alpha = .15, colour = "white")+ 
  geom_polygon(data = cnty_input_data, aes(x = long, y = lat, group = group,
                                           fill = value), alpha = .5, colour = "white") +
  geom_point(data = coca_port_list, aes(lon, lat),
             size = 2, color = "red") +
  ggrepel::geom_text_repel(data = coca_port_list, aes(x=lon, y=lat, label=port.nice),
                           direction = "both",
                           hjust = 1,
                           nudge_x = 10,
                           nudge_y = -1,
                           size = 2.7, #size of box?
                           segment.size = .3) +
  scale_fill_gradient(low = gmri.blue, high = gmri.gray) +
  coord_fixed(1.3, xlim = c(-74, -66.5), ylim =c(44.7, 40.65)) + 
  theme_map() + 
  theme(legend.position = "none",
        text = element_text(family = "Roboto Condensed"))
```

